Viscous Erosion in Porous Media

### 2018-01-18 docs/methods/erosion_methods.tex
NM

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation}
\label{s:formulation}
We start by defining the main variables used to model erosion.  We only
consider flows that are confined by a solid wall $\Gamma$ that encloses
$M$ eroding bodies.  The bodies are denoted as $\gamma_\ell$,
$\ell=1,\ldots,M$, and we write $\gamma = \gamma_1 \cup \cdots \cup
\gamma_M$.  Neglecting inertial forces, the dynamics of the fluid is
fully characterized by the position of the bodies $\xx_\ell(s,t) \in
\gamma_\ell$, where $s$ is the arclength and $t$ is time.  Given
$\xx_\ell(s,t)$, $\ell=1,\ldots,M$, derived variables include the fluid
velocity $\uu$, the pressure $p$, and the shear stress $\tau$.   On the
bounding wall, we prescribed a velocity $\UU(\xx,t)$.  Then, the
governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap
&&\mbox{conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{outer wall
velocity} \\
V_\nn = \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
&&\mbox{erosion model},
\end{split}
\end{equation}
where $V_\nn$ is the normal velocity of the interface and $\mu$ is the
fluid viscosity.  The outer wall is formed by rounding off the corners
of $[-3,3] \times [-1,1]$, and we enforce the Hagen-Poiseuille flow
\begin{align*}
  \uu(\xx) = (1-y^2,0)
\end{align*}
on $\Gamma$.  The boundary of the fluid domain is $\bd\Omega = \gamma_1
\cup \cdots \cup \gamma_M \cup \Gamma$.  To minimize the boundary
effects of the inflow and outflow, we only place bodies in the center
third, $[-1,1] \times [-1,1]$, of the fluid domain.
\begin{figure}[htpb]
  \centering
  \input{figs/schematic.tikz}
  \caption{\label{fig:schematic} A schematic for the governing
    equations.  A no-slip boundary condition is imposed on each body
    $\gamma_\ell$ whose unit normal points outward relative to the
    geometry.  On the outer geometry $\Gamma$, a pipe flow is imposed.
    The bodies are eroded at a rate that is proportional to the
    magnitude of the shear stress applied by the Stokesian fluid.  The
    bodies are constrained to the middle third of the channel that is
    located between the dashed lines.}
\end{figure}

@@

To model erosion, the normal velocity is prescribed to be
proportional to the absolute value of the shear stress.   Since the
constant of proportionality only sets the time scale, we let $V_\nn =
|\tau|$ as described in equation~\eqref{eqn:erosionModel}.
\begin{align}
  V{\nn} = \abs{\tau} \nn.
  \label{eqn:bodyEvolution}
\end{align}
For the sake of numerical stability, we modify this law in
Section~\ref{sec:thetaL}.

@@

Interestingly, a brief look at Fig.~\ref{shrink_intface}(a) reveals that the spacing between successive interfaces becomes larger as the body shrinks, suggesting that the overall shear stress is growing. Of course, such a relationship is consistent with the well-known 3D Stokes drag law, $\tau \sim \mu V / L$, since stress is inversely proportional to body scale $L$. This law must be modified in 2D, though. The modification is made non-trivial by subtleties associated with the Stokes paradox, but can be correctly deduced by noting that, as the body vanishes, its drag must be consistent with slender-body theory. For a slender body in flow perpendicular to its long-axis $W \gg L$, the drag is proportional to $\mu V W/ \log(W/L)$. This relationship implies that, for a 2D cross-section perpendicular to the long axis, the stress can be estimated by
\begin{equation}
\tau \sim \frac{\mu V}{L \log(W/L)}
\end{equation}
In interpreting our 2D simulations, we will treat $W$ as fixed (and must remember the requirement $W \gg L$).

@@

\subsubsection{The vanishing rate of area}

Then the rate of change of the body area is proportional to the shear
stress times the total arc length exposed
\begin{equation}
\dot{A} = \int \atauD{2} \, ds = \mean{\atauD{2}} L
\sim \frac{\mu V}{\log(W/L)}
%\sim - \mean{\abs{\tauD{2}}} L
\end{equation}
Solving gives (redo this calculation)
\begin{equation}
A (1 - \log A) = C (t_f - t)
\end{equation}
where $C$ is a constant.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation}
\label{s:formulation}
We start by defining the main variables used to model erosion.  We only
consider flows that are confined by a solid wall $\Gamma$ that encloses
$M$ eroding bodies.  The bodies are denoted as $\gamma_\ell$,
$\ell=1,\ldots,M$, and we write $\gamma = \gamma_1 \cup \cdots \cup
\gamma_M$.  Neglecting inertial forces, the dynamics of the fluid is
fully characterized by the position of the bodies $\xx_\ell(s,t) \in
\gamma_\ell$, where $s$ is the arclength and $t$ is time.  Given
$\xx_\ell(s,t)$, $\ell=1,\ldots,M$, derived variables include the fluid
velocity $\uu$, the pressure $p$, and the shear stress $\tau$.   On the
bounding wall, we prescribed a velocity $\UU(\xx,t)$.  Then, the
governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap
&&\mbox{conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{outer wall
velocity} \\
\Vn = \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
&&\mbox{erosion model},
\end{split}
\end{equation}
where $\Vn$ is the normal velocity of the interface and $\mu$ is the
fluid viscosity.  The outer wall is formed by rounding off the corners
of $[-3,3] \times [-1,1]$, and we enforce the Hagen-Poiseuille flow
\begin{align*}
  \UU(\xx) = \umax (1-y^2,0)
\end{align*}
on $\Gamma$.  H
The boundary of the fluid domain is $\bd\Omega = \gamma_1
\cup \cdots \cup \gamma_M \cup \Gamma$.  To minimize the boundary
effects of the inflow and outflow, we only place bodies in the center
third, $[-1,1] \times [-1,1]$, of the fluid domain.
\begin{figure}[htpb]
  \centering
  \input{figs/schematic.tikz}
  \caption{\label{fig:schematic} A schematic for the governing
    equations.  A no-slip boundary condition is imposed on each body
    $\gamma_\ell$ whose unit normal points outward relative to the
    geometry.  On the outer geometry $\Gamma$, a pipe flow is imposed.
    The bodies are eroded at a rate that is proportional to the
    magnitude of the shear stress applied by the Stokesian fluid.  The
    bodies are constrained to the middle third of the channel that is
    located between the dashed lines.}
\end{figure}

@@

To model erosion, the normal velocity is prescribed to be
proportional to the absolute value of the shear stress.   Since the
constant of proportionality only sets the time scale, we let $\Vn =
|\tau|$ as described in equation~\eqref{eqn:erosionModel}.
\begin{align}
  \Vn = \abs{\tau} \nn.
  \label{eqn:bodyEvolution}
\end{align}
For the sake of numerical stability, we modify this law in
Section~\ref{sec:thetaL}.

@@

Interestingly, a brief look at Fig.~\ref{shrink_intface}(a) reveals that the spacing between successive interfaces becomes larger as the body shrinks, suggesting that the overall shear stress is growing. Of course, such a relationship is consistent with the well-known 3D Stokes drag law, $\tau \sim \mu V / L$, since stress is inversely proportional to body scale $L$. This law must be modified in 2D, though. The modification is made non-trivial by subtleties associated with the Stokes paradox, but can be correctly deduced by noting that, as the body vanishes, its drag must be consistent with slender-body theory. For a slender body in flow perpendicular to its long-axis $W \gg L$, the drag is proportional to $\mu V W/ \log(W/L)$. This relationship implies that, for a 2D cross-section perpendicular to the long axis, the stress can be estimated by
\begin{equation}
\label{eq:stresslaw}
\tau \sim \frac{\mu V}{L \log(W/L)}
\end{equation}
In interpreting our 2D simulations, we will treat $W$ as fixed (and must remember the requirement $W \gg L$).

Now, the body's rate of area reduction is given by integrating the absolute shear stress around the boundary
\begin{equation}
\dot{A} = \int_{\gamma} \atau \, ds \, ,
\end{equation}
which gives the exact relationship
\begin{equation}
\dot{A} = \mean{\atau} L \, .
\end{equation}
Thus, the area-loss rate is given by the product of the mean shear $\mean{\atau}$ and the total perimeter $L$. Inserting the scaling law \eqref{eq:stresslaw} gives
 \begin{equation}
\dot{A} \sim \frac{\mu V}{\log(W/L)} \, .
\end{equation}
Noting that $A \sim L^2$ allows one to solve this ODE to get
\begin{equation}
A (1 - \log A) = C (t_f - t)
\end{equation}
This is an {\em implicit} relationship for how the area scales down in time during erosion.

\todo[inline]{Define $C$; also maybe need to use a different letter.}

@@
